The \dealii Library, Version 9.7

### 2025-08-02 9.7/paper.tex
Add one more note.
LH contributions.

\todo[inline]{Luca, want to say something here? If not, please delete
  the subsubsection.}

\subsubsection{PSBLAS, a XXXXX}

\todo[inline]{Complete section name}

\todo[inline]{Luca, want to say something here? If not, please delete
  the subsubsection.}
\todo[inline]{DRW: note that while we support configuring with PSBLAS we don't
  actually implement any interfaces to it yet (i.e., we have one test but no
  code in the library which uses \texttt{DEAL\_II\_WITH\_PSBLAS}).}

\subsubsection{\texttt{magic\_enum}, a XXXXX}

\todo[inline]{Complete section name}
\todo[inline]{Luca, want to say something here? If not, please delete
  the subsubsection.}

\subsubsection{CGAL, a computational geometry engine}

\todo[inline]{Do we want to say something here? If not, please delete
  the subsubsection.}

@@

\dealii{} currently requires a compiler that understands C++17, but
some additional features are enabled if a compiler supports the C++20
standard. C++20 also introduces ``modules'', a way by which software
packages can export their public interface without relying on the
mechanism of textual inclusion of header files that C++ has inherited from C. The
end goal of C++20 modules is that one can replace the use of long
lists of statements such as\\
\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h}\\
by a simple\\
\hspace*{1cm}\texttt{import dealii;}

@@

\dealii can interface with many other libraries:
\begin{multicols}{3}
  \begin{itemize}[leftmargin=4mm]
    \item ADOL-C \cite{griewank1996adolc}
    \item ArborX \cite{prokopenko2025}
    \item ARPACK \cite{lehoucq1998arpack}
    \item Assimp \cite{schulze2021assimp}
    \item BLAS and LAPACK \cite{anderson1999lapack}
    \item Boost \cite{boost-web-page}
    \item CGAL \cite{cgal:eb-24b}
    \item Gmsh \cite{geuzaine2009gmsh}
    \item GSL \cite{galassi2009gsl,gsl-web-page}
    \item Ginkgo \cite{anzt2020ginkgo,anzt2022ginkgo}
    \item HDF5 \cite{hdf5-web-page}
    \item Kokkos \cite{trott2022}
    \item Magic Enum \cite{magic-enum-web-page}
    \item METIS \cite{karypis1998metis}
    \item MUMPS \cite{amestoy2001mumps,amestoy2019mumps}
    \item muparser \cite{muparser-web-page}
    \item OpenCASCADE \cite{opencascade-web-page}
    \item p4est \cite{burstedde2011p4est}
    \item PETSc \cite{petsc-user-ref,petsc-web-page}
    \item ROL \cite{ridzal2014rol}
    \item ScaLAPACK \cite{blackford1997scalapack}
    \item SLEPc \cite{roman2023improvements}
    \item SUNDIALS \cite{gardner2022enabling}
    \item SymEngine \cite{symengine-web-page}
    \item Taskflow \cite{huang2021taskflow}
    \item TBB \cite{reinders2007tbb}
    \item Trilinos \cite{mayr2025trilinos,trilinos-web-page}
    \item UMFPACK \cite{davis2004umfpack}
    \item VTK \cite{vtk-web-page}
    \item zlib \cite{zlib-web-page}
  \end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use
interfaces to these libraries.

@@

>>>>

MUMPS (MUltifrontal Massively Parallel sparse direct Solver) is a widely-used software library designed for solving large sparse systems of linear equations efficiently on parallel architectures~\cite{amestoy2001mumps}.

In the 9.7 release, we have significantly enhanced integration with MUMPS,
without requiring it to be installed through PETSc or Trilinos. Improvements
include a simplified user interface for direct solvers, mimicking the existing
interface we provide for UMFPACK, and enhanced performance optimizations that
leverage recent MUMPS capabilities. Specifically, the deal.II wrappers now
transparently support parallel matrix factorization, out-of-core computations,
Block Low-Rank approximations, and GPU support.

\subsubsection{PSBLAS, a parallel sparse linear algebra library}

PSBLAS is a parallel sparse linear algebra library that provides basic operations for sparse matrices and vectors~\cite{Filippone2000psblas}.

The library is specifically designed to leverage modern
parallel computing architectures, including both CPU and GPU systems, making it well-suited for exascale
computing environments. Release 9.7 includes preliminary support for configuring with PSBLAS, allowing users of both frameworks to take advantage of its capabilities for specific linear algebra tasks. This support is still in its early stages, and the library is not yet fully integrated into the \dealii{} ecosystem.

\subsubsection{\texttt{magic\_enum}, a static reflection library for enums}

\texttt{magic\_enum} is a modern, header-only C++17 library designed to provide static reflection capabilities for enums without the need for macros or additional boilerplate code. It greatly simplifies working with enum types by enabling conversions between enum values and strings, integer values, and various other operations. Version 9.7 uses \texttt{magic\_enum} to provide type safe and efficient conversions between enum values and their string representations in the \texttt{ParameterHandler} class.

@@

\dealii{} currently requires a compiler that understands C++17, but
some additional features are enabled if a compiler supports the C++20
standard. C++20 also introduces ``modules'', a way by which software
packages can export their public interface without relying on the
mechanism of textual inclusion of header files that C++ has inherited from C. The
end goal of C++20 modules is that one can replace the use of long
lists of statements such as\\
\hspace*{1cm}\texttt{\#include <deal.II/grid/tria.h>}\\
by a simple\\
\hspace*{1cm}\texttt{import dealii;}

@@

\dealii can interface with many other libraries:
\begin{multicols}{3}
  \begin{itemize}[leftmargin=4mm]
    \item ADOL-C \cite{griewank1996adolc}
    \item ArborX \cite{prokopenko2025}
    \item ARPACK \cite{lehoucq1998arpack}
    \item Assimp \cite{schulze2021assimp}
    \item BLAS and LAPACK \cite{anderson1999lapack}
    \item Boost \cite{boost-web-page}
    \item CGAL \cite{cgal:eb-24b}
    \item Gmsh \cite{geuzaine2009gmsh}
    \item GSL \cite{galassi2009gsl,gsl-web-page}
    \item Ginkgo \cite{anzt2020ginkgo,anzt2022ginkgo}
    \item HDF5 \cite{hdf5-web-page}
    \item Kokkos \cite{trott2022}
    \item Magic Enum \cite{magic-enum-web-page}
    \item METIS \cite{karypis1998metis}
    \item MUMPS \cite{amestoy2001mumps,amestoy2019mumps}
    \item muparser \cite{muparser-web-page}
    \item OpenCASCADE \cite{opencascade-web-page}
    \item p4est \cite{burstedde2011p4est}
    \item PETSc \cite{petsc-user-ref,petsc-web-page}
    \item PSBLAS \cite{Filippone2000psblas}
    \item ROL \cite{ridzal2014rol}
    \item ScaLAPACK \cite{blackford1997scalapack}
    \item SLEPc \cite{roman2023improvements}
    \item SUNDIALS \cite{gardner2022enabling}
    \item SymEngine \cite{symengine-web-page}
    \item Taskflow \cite{huang2021taskflow}
    \item TBB \cite{reinders2007tbb}
    \item Trilinos \cite{mayr2025trilinos,trilinos-web-page}
    \item UMFPACK \cite{davis2004umfpack}
    \item VTK \cite{vtk-web-page}
    \item zlib \cite{zlib-web-page}
  \end{itemize}
\end{multicols}
Please consider citing the appropriate references if you use
interfaces to these libraries.

@@

This release contains patches written in and contributed from
Antarctica. As a consequence, \dealii{} now consists of work written
on all seven continents.
